A model-based method for detecting persistent cultural change using panel data[^thanks]

### 2021-02-08 ms.Rmd
conditional accept edits 2

This model constrains $\rho$ to be equal across waves. This implies two important assumptions. First, it assumes that the time between waves is equal. This is a reasonable assumption for the GSS, where surveys take place about two years apart. Across the three panels, the average time between waves 1 and 2 is between 695 and 764 days, while the average time between waves 2 and 3 is between 692 and 743 days. This might not be a reasonable assumption in other panels and can be loosened by adding a coefficient to $\rho$ for the time since last wave.

Second, constraining $\rho$ to be the same across waves assumes that active updating happens by a linear secular trend. While this is likely a valid assumption for many questions we explore in the GSS, there are questions where opinions likely change discontinuously and in durable ways around specific dates and events. In the GSS panels, these questions include confidence in various branches of the government, which might change as a result of elections or high-profile political events, and confidence in financial institutions, belief in the ability to find a job, and subjective socioeconomic status, which might change in response to the onset of the Great Recession.[^valid]

>>>>

This model constrains $\rho$ to be equal across waves. This implies that the rate of change between the two waves is equal, which comes with two implications. First, it assumes that the time between waves is equal. If waves are differently spaced, more time between waves is likely to equate to more active updating. This assumption is reasonable for the GSS panels, where surveys took place about two years apart.[^time] This might not be a reasonable assumption in other panels and can be loosened by adding a coefficient to $\rho$ for the time since last wave.

[^time]: Across the three panels, the average time between waves 1 and 2 is between 695 and 764 days, while the average time between waves 2 and 3 is between 692 and 743 days.

Second, constraining $\rho$ to be the same across waves implies that active updating happens by a linear secular trend. While this is likely a valid assumption for many questions we explore in the GSS, there are questions where opinions likely change discontinuously and in durable ways around specific dates and events. In the GSS panels, these questions include confidence in various branches of the government, which might change as a result of elections or high-profile political events, and confidence in financial institutions, belief in the ability to find a job, and subjective socioeconomic status, which might change in response to the onset of the Great Recession.[^valid]
